---
title: "ds_with_r"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
library(knitr)
```

Visualizing Alaska Air flights:
```{r}
alaska_air <- flights %>% filter(carrier == "AS")
```

A `glimpse(alaska_air)` call shows that there are a handful of flights each day, and that they typically all travel from `EWR` to `SEA`. Now, to visualize the Alaska Air flights using a scatter plot, drop in to `ggplot2#ggplot()`:
```{r}
ggplot(
  data=alaska_air, 
  mapping= aes(
    x = dep_delay,
    y = arr_delay
  )
) + geom_point(alpha=0.2)
```
This illustrates the _grammar of geometry_ discussed in the book (its a $100 text from Springer, and I don't need a copy). I am curious if this is the same grammar used by MBostok to derive D3? Regardless, its odd that `geom_point()` is added using a `+` at the end of the `ggplot()` definition. I expected this to be an argument rather than joining two different data structures together... 

Futher reading shows that the `geom_point` object is a distinct layer in the visualization. `ggplot2` works by constructing layers similar to how d3 does, so the `+` is actually adding a new layer on top of the basic struture defined via `ggplot()`. This suggests I could stack arbitrry layers on top of one another. I wonder if multiple ggplots can be stacked?

```{r}
ggplot(data=alaska_air, mapping=aes(x = arr_delay, y = dep_delay)) + 
geom_point() + 
ggplot(data=alaska_air, mapping=aes(x = air_time, y = distance)) + 
geom_bar()
```
Nope, can't stack multiple `ggplot()` objects with each other.

Considering the `weather` dataset, the delays could be caused by strong winds, heavy precipitation, and/or very low visibility. Combining two or more of these is expected to almost definitiely lead to delays. Having not yet joined the two datasets and examined the correlations, I'm unwilling to make a guess about the thresholds that define "strong", "heavy", or "very low", although I suspect that there is some sort of exponential function at work.

It is possible to alter the location of points in a visualization using `geom_jitter(height=<int>, width=<int>)`. This is a scatter visualization that randomly bumps points by some random amount in the range 0-> threshold, where threshold comes from the `height` and `width` arguments.
```{r}
early_jan_weather <- weather %>% 
  filter(month == 1 & day <= 15 & origin == "EWR")

ggplot(data=early_jan_weather, mapping=aes(x = time_hour, y = temp)) +
  geom_line()
```

Line graph of precipitation in Newark in ealry january
```{r}
ggplot(data=early_jan_weather, mapping = aes(x = time_hour, y = precip)) + 
  geom_line()
```

Experiment with plotting a distribution for a single numeric (measurement) variable:
```{r}
ggplot(data=weather, mapping = aes(x = temp, y = 0)) + geom_jitter(alpha = 0.1, height=0.1)
```
That's pretty interesting to illustrate the density - clusters between 30 & 80 - but it doesn't say much about how large the buckets are, only that there would be "a lot" of data in those buckets. A histogram is a better solution. It looks somewhat normally distributed around 60, although it also could be multi-modal around 45 and 65. That makes sense because temperatures will bottom in the winter and peak in the summer, but need to pass through both of those temperatures on their way in and out of the maxima. This is actually a really interesting principle to think about. Does it generalize and apply anywhere else in life?
```{r}
ggplot(data=weather, mapping = aes(x = temp)) + 
  geom_histogram(binwidth= 5, color = "pink") + 
  facet_wrap(~ month)
```
```{r}
ggplot(data=weather, mapping=aes(x = factor(month), y = temp)) + 
  geom_boxplot()
```
The box represents the IQR (inter-quartile range), bars (whiskers) are values between IQR & 1.5xIQR, and finally the single points are outliers.

For graphing the distribution of categorical variables, use `geom_bar()` or `geom_col()` depending on whether the data is pre-counted or not.
```{r}
ggplot(data=flights, mapping = aes(x = origin)) + 
  geom_bar() + 
  facet_wrap(~ carrier)
```
Another way to illustrate the above per-carrier flight numbers is with stacked and side-by-side bar charts. The default behavior is to stack bars, but for side-by-side, adding `position=dodge` to the geometry layer splits out the bars.
```{r}
ggplot(data=flights, mapping=aes(x=carrier, fill=origin)) +
  geom_bar()

ggplot(data=flights, mapping=aes(x=carrier,fill=origin)) + 
  geom_bar(position = position_dodge(preserve="single"))
```

