---
title: "R Notebook"
output: html_notebook
---

This notebook follow along with the early examples from https://moderndive.com/1-getting-started.html. The plan is to learn enough R to be productive and provide useful reviews at work.

Tidyverse libraries
```{r}
library(tidyverse)
library(nycflights13)
library(knitr)
```

Let's have a look at what's in the flights data
```{r}
flights
```
Timestamps, both actual and scheduled, delays between them, details about the in-air flight, as well as some categorical information like carrier, flight number, and start+end airports. This is a `tibble`, which is a specific type of dataframe in R that seems like it allows navigation in a worksheet or console like above.

```{r}
glimpse(flights)
```
Appears to transpose the data without any sampling or shuffling. This gives me a pretty good idea what the data looks like, as well as each column's datatype.
```{r}
kable(airlines)
```
`kable()` pretty-prints a dataframe. It seems to actually render markup when in an `rmd` file.

Individual fields (columns) of a dataframe can be accessed using the `$` operator. For example:

```{r}
airlines$na
```

Get help via `?flights`, or `?dplyr`. The `?` operator accesses the help docs for a given command/object.

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

