---
title: "Simple Linear Regression"
output: html_notebook
---
```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
```
In this notebook I explore simple linear regression. SLR is linear regression in a single variable, predicting a numerical output variable.

Let's start with some EDA to understand the shape of the data.
```{r}
evals_beauty <- evals %>%
  select(ID, score, bty_avg, age)
glimpse(evals_beauty)

evals_beauty %>%
   select(score, bty_avg, age) %>%
  skim()

evals_beauty %>%
  get_correlation(formula = score ~ bty_avg)

ggplot(evals_beauty, mapping = aes(x = score, y = bty_avg)) +
  geom_jitter() + 
  labs(x = "Score",
       y= "Beauty",
       title = "Score and Beauty in teachng")
```
You can add a fit line to visualizations using the technique of choice. The technique is provided as a string of the function name.
```{r}
ggplot(evals_beauty, mapping = aes(y = score, x = bty_avg)) +
  geom_jitter() + 
  labs(y = "Score",
       x= "Beauty",
       title = "Score and Beauty in teachng") +
  geom_smooth(method="lm")
```
The line plotted above is called a regression line. It illustrates the impact that a change in beauty has on score. In this case, the slope is quite low, as fitting the low positive correlation between beauty and score.To determine a predicted score based on the beauty_avg, let's run a linear regresson.
```{r}
model <- lm(score ~ bty_avg, evals)
get_regression_table(model)
```
The regression table above shows that the majority of the score comes from the intercept rather than the effect of beauty. That fits with the low correlation found earlier.The `lm` function trains a linear model based on the provied forumla. 

Open question: How to train a model in multiple variables?

Same analysis but for age:
```{r}
age_model <- lm(score ~ age, evals)
get_regression_table(age_model)
get_regression_points(model)
```
Flipping over to look at regression with a single categorical variable. Open question: How to mix categorical and numerical variables together in a regression?

```{r}
library(gapminder)

gapminder2007 <- gapminder %>%
  filter(year == 2007) %>%
  select(- year)

glimpse(gapminder2007)
gapminder2007 %>% sample_n(10)
skim(gapminder2007)

ggplot(gapminder2007, mapping=aes(x=lifeExp)) +
  geom_histogram(binwidth = 5) +
  labs(title="Life expectancies") +
  facet_wrap(~ continent)


ggplot(gapminder2007, mapping=aes(x=continent, y= lifeExp)) +
  geom_boxplot()

ggplot(gapminder2007, mapping=aes(x=continent, y=gdpPercap)) + 
  geom_boxplot()
```

